---
title: "Team Capacity Management Dashboard"
# author: "Michalis Zouvelos"
# date: "May 2022"
output: 
  flexdashboard::flex_dashboard:
    social: ["menu"]
    source_code: embed
    vertical_layout: fill
    theme: flatly
runtime: shiny
---

``` {js}
// Inverse color of navigation bar.
$('.navbar-inverse').removeClass('navbar-inverse').addClass('navbar-default');
```


```{r setup, include=FALSE}
# Loading the pacman library
if (!require(pacman)) install.packages('pacman')
library(pacman)
# Importing the required Packages using pacman
pacman::p_load(tidyverse, vroom, flexdashboard, lubridate, shiny, plyr, tidyr, DT, htmlwidgets, shiny, shinydashboard, highcharter, DT, viridis, shinyBS, shinyjs, withr, treemap, WDI, shinycssloaders, timevis, bslib, janitor, tibble, echarts4r)
```

```{r read, include = FALSE}
# reading the combined Capacity excel file

capacity <- readxl::read_excel("data.xlsx")


# renaming columns
capacity <- capacity %>%
  dplyr::rename(Employee = `Employee Name`) %>% 
  dplyr::rename(Contract_end = `Contract End`) %>% 
  dplyr::rename(Project = `Project Name`) %>% 
  dplyr::rename(Proj_Status = Status) %>% 
  dplyr::rename(Progress = `Project Progress`)

cols.num <- c(8:59)
capacity[cols.num] <- sapply(capacity[cols.num],as.character)


# Transformation of the wide data set format to a long format to track week series
capacity <- pivot_longer(capacity,cols = Week_1:Week_52, names_to = "Week", names_prefix = "Week_", values_to = "value") # Pivot weekly values to Long Format

capacity$Progress <- as.numeric(capacity$Progress) # Turning progress of the project to numeric
capacity$value <- as.numeric(capacity$value) # Turning value column into numeric from factor
capacity$value[is.na(capacity$value)] <- 0 # Replacing missing values with 0

capacity$Week <- as.numeric(capacity$Week) # Turning week column into numeric

# capacity$Contract_end <- as.numeric(capacity$Contract_end) # Turning Contract end date to numeric
# capacity$Contract_end <- excel_numeric_to_date(as.numeric(as.character(capacity$Contract_end)), date_system = "modern") # Turning Contract date to proper date, modern style 

# Getting Date extracted from Week number
capacity <- capacity %>% 
  mutate(week_date = as.Date(paste(2022, capacity$Week, 1, sep = "-"), "%Y-%U-%u"))

capacity$full_date <- capacity$week_date

capacity$week_date <- format(ymd(capacity$week_date), "%d-%m")

capacity$day <- sapply(capacity$week_date, FUN = function(x) {strsplit(x, split = '[-]')[[1]][1]})
capacity$month <- sapply(capacity$week_date, FUN = function(x) {strsplit(x, split = '[-]')[[1]][2]})

capacity$day <- as.numeric(capacity$day)
capacity$month <- as.numeric(capacity$month)
capacity$month <- month.abb[capacity$month]

capacity$week_full <- paste(capacity$month, capacity$day, sep = "-")

# Make dataframe of week numbers and date ranges
week_df <- capacity %>% 
  distinct(Week, week_full)

# Custom function to extract the dates for every friday in desired date range
pick.wkday <- function(selday,start,end) {
  fwd.7 <- start + 0:6
  first.day <- fwd.7[as.numeric(format(fwd.7,"%w"))==selday]
  seq.Date(first.day,end,by="week")
}

# Extracting all Dates for Fridays in 2022
week_df$friday_date <- pick.wkday(5,as.Date("2022-01-01"), as.Date("2022-12-31"))

# keep only the day from the friday date
week_df$friday_date <- as.character(week_df$friday_date)
week_df$friday <- sapply(week_df$friday_date, FUN = function(x) {strsplit(x, split = '[-]')[[1]][3]})
week_df$friday <- as.numeric(week_df$friday)
capacity$week_full <- paste(capacity$month, capacity$day, sep = "-")
week_df$week_full <- paste(week_df$week_full, week_df$friday, sep = "-")

capacity$value <- capacity$value*100 # Turning values to represent percentages
```

```{r include=FALSE}
# Plot style functions

# valueBox style custom function
VB_style <- function(msg = 'Message', style="font-size: 100%;"){
   tags$p( msg , style = style )
}

# Create a custom theme for the highchart plots 
custom_theme <- hc_theme(
  colors = c('#5CACEE', 'green', 'red'),
  chart = list(
         backgroundColor = '#FAFAFA', 
         plotBorderColor = "black"),
  xAxis = list(
         gridLineColor = "E5E5E5", 
         labels = list(style = list(color = "#333333")), 
         lineColor = "#E5E5E5", 
         minorGridLineColor = "#E5E5E5", 
         tickColor = "#E5E5E5", 
         title = list(style = list(color = "#333333"))), 
  yAxis = list(
         gridLineColor = "#E5E5E5", 
         labels = list(style = list(color = "#333333")), 
         lineColor = "#E5E5E5", 
         minorGridLineColor = "#E5E5E5", 
         tickColor = "#E5E5E5", 
         tickWidth = 1, 
         title = list(style = list(color = "#333333"))),   
  title = list(style = list(color = '#333333', fontFamily = "Lato")),
  subtitle = list(style = list(color = '#666666', fontFamily = "Lato")),
  legend = list(
         itemStyle = list(color = "#333333"), 
         itemHoverStyle = list(color = "#FFF"), 
         itemHiddenStyle = list(color = "#606063")), 
  credits = list(style = list(color = "#666")),
  itemHoverStyle = list(color = 'gray'))
```

<style>
.colored {
  background-color: #FAFAFA;
}
</style>

Department Overview {data-orientation=rows data-icon="fa-sitemap"}
=======================================================================

Column {.sidebar data-width=250}
-----------------------------------------------------------------------
```{r}
type_choice <- capacity$Type %>% 
  na.omit() %>% 
  sort()

#  Importing slider filter for Team Member Type
selectInput(inputId = "v_type", 
           label = h3('Pick Type here:'), unique(type_choice), selected = "")

week_choice <- capacity$Week %>% 
  na.omit() %>% 
  sort()

# Importing slider filter for Week - default selection as current week
selectInput(inputId = "v_week", 
           label = h3('Pick Week here:'), unique(week_choice), selected = strftime(Sys.Date(), format = "%V"))



# Display Datatable to be able to scroll through the weeks and see date range of each week of the year

week_df %>%
    distinct(Week, week_full) %>% 
  DT::datatable(rownames = FALSE, colnames = c('Week', 'Date Range'),
                class = 'cell-border stripe',
                options = list(autoWidth = TRUE, scrollY = 600, scrollX = FALSE, paging = FALSE, dom = 't', ordering = FALSE,
                columnDefs = list(list(className = 'dt-left', targets = "_all"),list(width ="40px", targets = 0),list(width ="115", targets = 1))))


```

Row {data-height=500}
-----------------------------------------------------------------------

### Team Total Outlook by Type (filter Contractor or Internal) {.no-padding}

```{r}
# data frame
overview_type <- capacity %>%
  select(Week, value, Type, Employee, week_full) %>%
  na.omit() %>% 
  dplyr::group_by(Week, Employee, Type, week_full) %>% 
  dplyr::summarise(value = sum(value)) %>% 
  arrange(Week) 
```

```{r}
# Highchart plot visualizing all same type team members total weekly involvement 
output$plot_type <- renderHighchart({
highchart() %>%
                         #    hc_exporting(enabled = TRUE, formAttributes = list(target = "_blank")) %>%
                             hc_add_series( data =  overview_type %>% filter(Type %in% input$v_type) ,
                                            mapping = hcaes(  x = Week, y = value, group = Employee ),
                                            type = 'line',
                                            marker = list(symbol = 'circle') ,
                                            visible = c(T,rep(F,length(10)-1))
                             ) %>%
                          #   hc_xAxis( categories = c( unique( p$Week) ) ) %>%
                             hc_xAxis( title = list(text = "Weeks")) %>%
                             hc_yAxis( title = list(text = "Involvement per Week (Employee)"), 
                                       allowDecimals = TRUE,labels = list(format = "{value}%")) %>%
                             hc_plotOptions(line = list(
                                dataLabels = list(enabled = F),
                                enableMouseTracking = T)
                             )%>%
                             hc_tooltip(table = TRUE,
                                        sort = TRUE,
                                        pointFormat = paste0( '<br> <span style="color:{point.color}">\u25CF</span>',
                                                              "{series.name}: <b>{point.y}%</b> "),
                                        headerFormat = '<span style="font-size: 13px"><b>Week {point.key}</b></span>'
                             ) %>%
                             hc_legend( layout = 'vertical', align = 'right', verticalAlign = 'top', floating = T, x = -50, y = -15 )
})

highchartOutput('plot_type')

```


Row {data-height=500}
-----------------------------------------------------------------------

### {.no-padding}


```{r}
output$plot_weekly <- renderHighchart({
weekly_capacity_internal <- capacity %>%
  dplyr::filter(Type == "Internal") %>% 
  dplyr::select(Employee, value, Week) %>% 
  dplyr::group_by(Employee, Week) %>% 
  dplyr::summarise(value = sum(value)) %>% 
  dplyr::filter(Week %in% input$v_week) %>% 
  dplyr::arrange(-value)
  
weekly_capacity_internal %>% 
 hchart(type = 'column', options3d = list(enabled = TRUE, alpha = 15, beta = 15), hcaes(x = reorder(Employee,value), y = value), color = "#d3d3d3", name = "Weekly Capacity") %>% 
 hc_yAxis(title = list(text = "Weekly Capacity"),labels = list(format = "{value}%")) %>%  
 hc_xAxis(title = list(text = "Resource"), labels = list(rotation=-45)) %>% 
 hc_title(text = "Weekly <b>Internal Employees</b> capacity", style = list(color = "#000000", useHTML = TRUE), labels = list(format = "{value}%")) %>% 
 hc_plotOptions(series = list(stacking = FALSE, allowPointSelect = TRUE), column = list(
    colorByPoint = FALSE,
    pointPadding = 0,
    groupPadding = 0.05,
    states = list(
      hover = list(
        enabled = FALSE
      )
    ),
    animation = FALSE,
    dataLabels = list(
      align = "center",
      enabled = TRUE,                                   
      formatter = JS(paste0("function() {                              
                           var max = this.series.yAxis.max,
                           color = this.y ? '#000000' : '#000000';
                           return '<span style=\"color: ' + color + '\">' + this.y + '%</span>' + '<span style=\"color: red\"> ';
    }")),
      style = list(color = "#0D538D", textShadow = "0"),
      shadow = FALSE,
      x = 1,
      y = 1
    ))
  ) %>% 
  hc_chart(events = list(load = JS("function() {
        let chart = this,
          x;

        chart.series[0].points.forEach(p => {
          if (p.value > 100) {
            x = p.x;
            p.graphic.element.style.fill = 'red';
          }
        })
        chart.series[0].points.forEach(p => {
          if (p.value <= 100 && p.value > 80) {
            x = p.x;
            p.graphic.element.style.fill = 'yellow';
          }
        })
        chart.series[0].points.forEach(p => {
          if (p.value <= 80) {
            x = p.x;
            p.graphic.element.style.fill = 'green';
          }
        })
      }
  ")))
  
})

highchartOutput('plot_weekly')
```

### {.no-padding}

```{r}
output$coolplot_weekly2 <- renderHighchart({
weekly_capacity_external <- capacity %>%
  dplyr::filter(Type == "Contractor") %>% 
  dplyr::select(Employee, value, Week) %>% 
  dplyr::group_by(Employee, Week) %>% 
  dplyr::summarise(value = sum(value)) %>% 
  dplyr::filter(Week %in% input$v_week) %>% 
  dplyr::arrange(-value)
  
weekly_capacity_external %>% 
 hchart(type = 'column', options3d = list(enabled = TRUE, alpha = 15, beta = 15), hcaes(x = reorder(Employee,value), y = value), color = "#d3d3d3", name = "Weekly Capacity") %>% 
 hc_yAxis(title = list(text = "Weekly Capacity"), labels = list(format = "{value}%")) %>%  
 hc_xAxis(title = list(text = "Resource"), labels = list(rotation=-45)) %>% 
 hc_title(text = "Weekly <b>External Contractors</b> capacity", style = list(color = "#000000", useHTML = TRUE)) %>% 
 hc_plotOptions(series = list(stacking = FALSE, allowPointSelect = TRUE), column = list(
    colorByPoint = FALSE,
    pointPadding = 0,
    groupPadding = 0.05,
    states = list(
      hover = list(
        enabled = FALSE
      )
    ),
    animation = FALSE,
    dataLabels = list(
      align = "center",
      enabled = TRUE,                                    
      formatter = JS(paste0("function() {                              
                           var max = this.series.yAxis.max,
                           color = this.y ? '#000000' : '#000000';
                           return '<span style=\"color: ' + color + '\">' + this.y + '%</span>' + '<span style=\"color: red\"> ';
    }")),
      style = list(color = "#0D538D", textShadow = "0"),
      shadow = FALSE,
      x = 1,
      y = 1
    ))
  ) %>% 
  hc_chart(events = list(load = JS("function() {
        let chart = this,
          x;

        chart.series[0].points.forEach(p => {
          if (p.value > 100) {
            x = p.x;
            p.graphic.element.style.fill = 'red';
          }
        })
        chart.series[0].points.forEach(p => {
          if (p.value <= 100 && p.value > 80) {
            x = p.x;
            p.graphic.element.style.fill = 'yellow';
          }
        })
        chart.series[0].points.forEach(p => {
          if (p.value <= 80) {
            x = p.x;
            p.graphic.element.style.fill = 'green';
          }
        })
      }
  ")))

  
})

highchartOutput('coolplot_weekly2')
```


Department Projects {data-icon="fa-calculator"}
=======================================================================

Column {.sidebar data-width=250}
-----------------------------------------------------------------------

```{r}
project_choice <- capacity$Project %>% 
  na.omit() %>% 
  sort()

# Importing dropdown filter for Projects
selectInput(inputId = "v_project", 
           label = h3('Pick Project here:'), unique(project_choice), selected = "")

# Display Datatable to be able to scroll through the weeks

week_df %>%
    distinct(Week, week_full) %>% 
  DT::datatable(rownames = FALSE, colnames = c('Week', 'Date Range'),
                class = 'cell-border stripe',
                options = list(autoWidth = TRUE, scrollY = 600, scrollX = FALSE, paging = FALSE, dom = 't', ordering = FALSE,
                columnDefs = list(list(className = 'dt-left', targets = "_all"),list(width ="40px", targets = 0),list(width ="115", targets = 1))))

```


Column {.tabset .tabset-fade data-width=750 data-height = 500 .colored }
-----------------------------------------------------------------------

### {.no-padding}

```{r}
t_proj <- capacity %>%
  select(Week, value, Project, week_full) %>%
  na.omit() %>% 
  dplyr::group_by(Week, Project, week_full) %>% 
  dplyr::summarise(value = sum(value)/100) %>% 
  arrange(Week) 

```


```{r}
# Highchart plot for tracking weekly total involvement of the team per Project
output$plot4 <- renderHighchart({
# Create line plot.
highchart() %>% 
    hc_add_series(t_proj %>% 
                    filter(Project %in% input$v_project)
                  , hcaes(x = Week, y = value, color = value), type = "line") %>%
  #  hc_tooltip(crosshairs = TRUE, borderWidth = 1.5, pointFormat = paste("Week: <b>{point.x}</b> <br> Value: <b>{point.y}</b>", "<br>Resource: <b>{Resource}</b>",
  #                              headerFormat = '<span style="font-size: 13px"><b>Week {point.key}</b></span>'  )) %>%
    hc_tooltip(table = TRUE,
                        sort = TRUE,
                        pointFormat = paste0(" ~ <b>{point.week_full}</b><br>Involvement: <b>{point.y}</b> "),
                        headerFormat = '<span style="font-size: 13px"><b>Week {point.key}</b></span>') %>% 
    hc_title(text = "Project Total Capacity by week") %>% 
    hc_subtitle(text = "2022") %>%
    hc_xAxis(title = list(text = "Weeks")) %>%
    hc_yAxis(title = list(text = "Involvement per week (Total)"),
             allowDecimals = TRUE, 
             plotLines = list(list(
                    color = "black", width = 1, dashStyle = "Dash", 
                    label = list(text = "", align = "left",
                                 style = list(color = "black"))))) %>%
    hc_legend(enabled = FALSE)  %>% 
    hc_add_theme(custom_theme)  %>% 
  hc_chart(events = list(load = JS("function() {
        let chart = this,
          x;

        chart.series[0].points.forEach(p => {
          if (p.value > 1) {
            x = p.x;
            p.graphic.element.style.fill = '#FF0000';
          }
        })
        chart.series[0].points.forEach(p => {
          if (p.value > 0.8 && p.value <= 1) {
            x = p.x;
            p.graphic.element.style.fill = '#EACE09';
          }
        })
        chart.series[0].points.forEach(p => {
          if (p.value <= 0.8) {
            x = p.x;
            p.graphic.element.style.fill = '#228b22';
          }
        })
        

      }
  ")))
})

highchartOutput('plot4')


```


### Project Resources {.no-padding}

```{r, include = FALSE}
p_proj <- capacity %>%
  select(Week, value, Employee, Project, week_full) %>%
  na.omit() %>% 
  dplyr::group_by(Week, Employee, Project, week_full) %>% 
  dplyr::summarise(value = sum(value)) %>% 
  arrange(Week)

```

```{r}
# Highchart plot for tracking involvement of each team member at selected project
output$plot5 <- renderHighchart({
highchart() %>%
                         #    hc_exporting(enabled = TRUE, formAttributes = list(target = "_blank")) %>%
                             hc_add_series( data =  p_proj %>% filter(Project %in% input$v_project) ,
                                            mapping = hcaes(  x = Week, y = value, group = Employee ),
                                            type = 'line',
                                            marker = list(symbol = 'circle') ,
                                            visible = c(T,rep(F,length(10)-1))
                             ) %>%
                          #   hc_xAxis( categories = c( unique( p$Week) ) ) %>%
                             hc_xAxis( title = list(text = "Weeks")) %>%
                             hc_yAxis( title = list(text = "Involvement % per Week (Employee)", plotLines = list(list(
                    color = "black", width = 1, dashStyle = "Dash", 
                    value = 80, labels = list(format = "{value}%"),
                    label = list(text = "Target = 80%", align = "left",
                                 style = list(color = "black")))), 
                                       allowDecimals = TRUE)) %>%
                             hc_plotOptions(line = list(
                                dataLabels = list(enabled = F),
                                #stacking = "normal",
                                enableMouseTracking = T)
                             )%>%
                             hc_tooltip(table = TRUE,
                                        sort = TRUE,
                                        pointFormat = paste0( '<br> <span style="color:{point.color}">\u25CF</span>',
                                                              "{series.name}: <b>{point.y}%</b> "),
                                        headerFormat = '<span style="font-size: 13px"><b>Week {point.key}</b></span>'
                             ) %>%
                             hc_legend( layout = 'vertical', align = 'right', verticalAlign = 'top', floating = T, x = -50, y = -15 )
})

highchartOutput('plot5')

```



Column {data-width=150}
-----------------------------------------------------------------------

### Project Status {.value-box}

```{r}
# Project Status value box
output$box_type2 <- flexdashboard::renderValueBox({
k2 <- capacity %>% 
  mutate(reactive_color = case_when(Proj_Status == "In progress" ~ "green",
                                    Proj_Status == "Cancelled" ~ "red",
                                    Proj_Status == "Paused" ~ "orange",
                                    Proj_Status == "Planned" ~ "grey")) %>% 
  select(Project, Proj_Status, reactive_color) %>% 
  filter(Project %in% input$v_project) %>% 
  na.omit()

flexdashboard::valueBox(
           VB_style( paste0(format(unique(k2$Proj_Status),big.mark=','), " "), "font-size: 100%; font-weight: bold;"  ),
            "",  
            color = unique(k2$reactive_color)
         )
})

flexdashboard::valueBoxOutput("box_type2")
```


### Business Department {.value-box}

```{r}
# Business Department of the Project focus
output$box_type1 <- flexdashboard::renderValueBox({
k1 <- capacity %>% 
  select(Project, Area) %>% 
  filter(Project %in% input$v_project) %>% 
  na.omit

flexdashboard::valueBox(
           VB_style( paste0(format(unique(k1$Area),big.mark=','), " "), "font-size: 100%; font-weight: bold;"  ),
            "",
         )
})

flexdashboard::valueBoxOutput("box_type1")
```

### 

```{r}
# Cool echart package plot to show percentage of project progression
renderEcharts4r({
progress <- capacity %>%
  select(Project, Progress) %>% 
  dplyr::filter(Project %in% input$v_project) %>% 
  na.omit

progression <- unique(progress$Progress)


  e_charts() |>
    e_gauge(progression, "PERCENT")|> 
    e_title("Project Progress" ,subtext = "% of Project Completion")
  })
```

### 

```{r}
# Pie chart by the type of employee in each project
output$plot6 <- renderHighchart({
  
pie_df <- capacity %>% 
  dplyr::group_by(Project, Type, Employee) %>%
  dplyr::summarise() %>%
  dplyr::select(Project, Type, Employee) %>% 
  na.omit() %>% 
  filter(Project %in% input$v_project)

hchart(pie_df$Type,	"pie", color = "#354e5a", name = "Project Status") %>% 
  hc_tooltip(borderWidth = 1.5, headerFormat = "<b>{point.key}</b><br>", pointFormat = paste("Percentage: <b> {point.percentage:.2f}%</b> <br> Amount by Type: <b>{point.y}</b>")) %>%
  hc_title(text = "<b>Resource Type Distribution</b>")  %>% 
   hc_plotOptions(pie = list(dataLabels = list(distance = 5, 
                            style = list(fontSize = 10)), 
                            size = 150)) 
})
highchartOutput('plot6')

```


Department Employees {data-icon="fa-address-card"}
=======================================================================

Column {.sidebar data-width=250}
-----------------------------------------------------------------------

```{r}
employee_choice <- capacity$Employee %>% 
  na.omit() %>% 
  sort()

selectInput(inputId = "v_employee", 
           label = h3('Pick Employee here:'), unique(employee_choice), selected = "")

# Display Datatable to be able to scroll through the weeks
week_df %>%
    distinct(Week, week_full) %>% 
  DT::datatable(rownames = FALSE, colnames = c('Week', 'Date Range'),
                class = 'cell-border stripe',
                options = list(autoWidth = TRUE, scrollY = 600, scrollX = FALSE, paging = FALSE, dom = 't', ordering = FALSE,
                columnDefs = list(list(className = 'dt-left', targets = "_all"),list(width ="40px", targets = 0),list(width ="115", targets = 1))))
```

Column {.tabset .tabset-fade data-width=750 data-height = 500 .colored }
-----------------------------------------------------------------------

### {.no-padding}

```{r}
df_t <- capacity %>%
  select(Week, value, Employee, week_full) %>%
  na.omit() %>% 
  dplyr::group_by(Week, Employee, week_full) %>% 
  dplyr::summarise(value = sum(value)) %>% 
  arrange(Week)
```


```{r}

output$plot <- renderHighchart({
# Create line plot.
highchart() %>% 
    hc_add_series(df_t %>% 
                    filter(Employee %in% input$v_employee)
                  , hcaes(x = Week, y = value, color = value), type = "line") %>%
    hc_tooltip(table = TRUE,
                        sort = TRUE,
                        pointFormat = paste0(" ~ <b>{point.week_full}<br>Involvement: <b>{point.y}%</b> "),
                        headerFormat = '<span style="font-size: 13px"><b>Week {point.key}</b></span>') %>% 
    hc_title(text = "Employee Total % Weekly Capacity") %>% 
    hc_subtitle(text = "2022") %>%
    hc_xAxis(title = list(text = "Weeks")) %>%
    hc_yAxis(title = list(text = "Involvement % per week (Total)"),
             allowDecimals = TRUE, labels = list(format = "{value}%"),
             plotLines = list(list(
                    color = "black", width = 1, dashStyle = "Dash", 
                    value = 80,
                    label = list(text = "Target = 80%", align = "left",
                                 style = list(color = "black"))))) %>%
    hc_legend(enabled = FALSE)  %>% 
    hc_add_theme(custom_theme)  %>% 
  hc_chart(events = list(load = JS("function() {
        let chart = this,
          x;

        chart.series[0].points.forEach(p => {
          if (p.value > 100) {
            x = p.x;
            p.graphic.element.style.fill = '#FF0000';
          }
        })
        chart.series[0].points.forEach(p => {
          if (p.value > 80 && p.value <= 100) {
            x = p.x;
            p.graphic.element.style.fill = '#EACE09';
          }
        })
        chart.series[0].points.forEach(p => {
          if (p.value <= 80) {
            x = p.x;
            p.graphic.element.style.fill = '#228b22';
          }
        })
        

      }
  ")))
})

highchartOutput('plot')


```


### Resource Projects Involvement {.no-padding}

```{r, include = FALSE}
df_p <- capacity %>%
  select(Week, value, Employee, Project) %>%
  na.omit() %>% 
  dplyr::group_by(Week, Employee, Project) %>% 
  dplyr::summarise(value = sum(value)) %>% 
  arrange(Week)
```

```{r}
output$plot2 <- renderHighchart({
highchart() %>%
                            # hc_exporting(enabled = TRUE, formAttributes = list(target = "_blank")) %>%
                             hc_add_series( data =  df_p %>% filter(Employee %in% input$v_employee) ,
                                            mapping = hcaes(  x = Week, y = value, group = Project ),
                                            type = 'line',
                                            marker = list(symbol = 'circle') ,
                                            visible = c(T,rep(F,length(10)-1))
                             ) %>%
                          #   hc_xAxis( categories = c( unique( p$Week) ) ) %>%
                             hc_xAxis( title = list(text = "Weeks")) %>%
                             hc_yAxis( title = list(text = "Weekly % Involvement per Project"), 
                                       allowDecimals = TRUE, labels = list(format = "{value}%")) %>%
                             hc_plotOptions(line = list(
                                dataLabels = list(enabled = F),
                                #stacking = "normal",
                                enableMouseTracking = T)
                             )%>%
                             hc_tooltip(table = TRUE,
                                        sort = TRUE,
                                        pointFormat = paste0( '<br> <span style="color:{point.color}">\u25CF</span>',
                                                              " {series.name}: <b>{point.y}%</b> "),
                                        headerFormat = '<span style="font-size: 13px"><b>Week {point.key}</b></span>'
                             ) %>%
                             hc_legend( layout = 'vertical', align = 'right', verticalAlign = 'top', floating = T, x = -50, y = -15 )
})

highchartOutput('plot2')

```



Column {data-width=150}
-----------------------------------------------------------------------

### Employee Type {.value-box}

```{r}

output$box_type <- flexdashboard::renderValueBox({
df_k <- capacity %>% 
  mutate(reactive_color = case_when(Type == "Internal" ~ "green",
                                    Type == "Contractor" ~ "red")) %>% 
  select(Employee, Type, reactive_color) %>% 
  filter(Employee %in% input$v_employee) %>% 
  na.omit

flexdashboard::valueBox(
           VB_style( paste0(format(unique(df_k$Type),big.mark=','), " "), "font-size: 100%; font-weight: bold;"  ),
            "" ,  
            color = unique(df_k$reactive_color)
         )
})

flexdashboard::valueBoxOutput("box_type")
```

### Contract End Date {.value-box}

```{r}
output$box_type_end <- flexdashboard::renderValueBox({
contract <- capacity %>% 
  select(Employee, Contract_end) %>% 
  filter(Employee %in% input$v_employee) %>% 
  na.omit

flexdashboard::valueBox(
           VB_style( paste0(format(unique(contract$Contract_end),big.mark=','), " "), "font-size: 100%; font-weight: bold;"  ),
            "", 
         )
})

flexdashboard::valueBoxOutput("box_type_end")

```

### 

```{r}
# Cool echart package plot to show percentage of utilization of each emplyoee to date
renderEcharts4r({
utilization <- capacity %>%
  dplyr::filter(Employee %in% input$v_employee) %>% 
  dplyr::select(Employee, value, Week) %>% 
  dplyr::group_by(Employee, Week) %>% 
  dplyr::summarise(value = sum(value)/100) %>% 
  dplyr::filter(Week <= strftime(Sys.Date(), format = "%V")) # get current week, and filter the data values up to this week

performance <- round(mean(utilization$value),2)

liquid <- data.frame(val = c(performance, performance-0.1, performance-0.2))
    liquid |>
      e_charts() |>
      e_liquid(val)|> 
  e_title("2022 Employee Utilization %" ,subtext = "Up to current week")
  })
```


### 

```{r}
output$plot3 <- renderHighchart({
df_pie2 <- capacity %>% 
  dplyr::group_by(Project, Proj_Status, Employee) %>%
  dplyr::summarise() %>%
  dplyr::select(Project, Proj_Status, Employee) %>% 
  na.omit() %>% 
  filter(Employee %in% input$v_employee)

hchart(df_pie2$Proj_Status,	"pie", color = "#354e5a", name = "Project Status") %>% 
  hc_tooltip(borderWidth = 1.5, headerFormat = "<b>{point.key}</b><br>", pointFormat = paste("Percentage: <b> {point.percentage:.2f}%</b> <br> Number of Projects: <b>{point.y}</b>")) %>%
  hc_title(text = "<b>Projects by Status</b>")  %>% 
   hc_plotOptions(pie = list(dataLabels = list(distance = 5, 
                            style = list(fontSize = 10)), 
                            size = 150)) 
})
highchartOutput('plot3')
```

About {data-icon="fa-info-circle"}
=======================================================================

Column {data-width=600}
-----------------------------------------------------------------------

### { .colored }


**Thank you for checking out my dashboard**


If you want to see more of my past and future projects you can find them on my personal website portfolio or on my GitHub, links below:


* [Michalis Zouvelos - Personal Website/Portfolio - Blog](https://mzouvelos.github.io/)

* [GitHub](https://github.com/mzouvelos)


<br> 


**About the dashboard**


The dashboard's **purpose** is to **solve the common problem** that many managers and companies face when trying to distribute their resources and plan their future capacity. 
Different departments have numerous projects and a plethora of resources of different types. Thus, allocating these resources correctly and tracking each project's involvement 
and individual involvement is crucial.

Instead of spending hours of checking project and team capacity in a spreadsheet, an interactive dashboard is visually pleasing. It offers a more convenient and easy way of 
tracking these metrics. The dashboard has three main views-pages:


- **A Department Overview view:** Where you can filter by Type of Employee to get an overall view of their total weekly capacity by percentage and then filter by Week to 
reallocate the workload between Employees and plan the capacity for future weeks.


- **A Department Projects view:** Where you can filter by a project to see the total capacity allocated on the project, each Resource who is involved in the selected project, 
and their weekly involvement in the project. Additionally, you can track the Project Status, the Business Area, the Project Progress, and the Employee Distribution by Resource 
Type.

- **A Department Employees view:** Where you can filter by Employee to check their total weekly capacity, the projects they are involved in, and how much they are involved in 
each of the projects weekly. On top of that, there is information about their employment type, either internal or contractor, which would then include a contract end date and 
their total utilization % up to the current date week.
</p>

The dashboard was created with [**RMarkdown**](https://rmarkdown.rstudio.com/) using the [**Flexdashboard**](https://pkgs.rstudio.com/flexdashboard/index.html) framework and 
adding [**Shiny**](https://shiny.rstudio.com/) components to enable filters. Most plots were created with the Highcharts package, which offers interactive charts that you can 
adjust further using Javascript. The code is available by clicking on the Source Code button.


The data used to create the dashboard are an example of **fictional data** (featuring **Star Wars characters** as employees of the Galaxy Empire Co.) generated by me to serve 
the purpose of creating the dashboard; thus, cleaning and preprocessing are limited in the code.



